---
title: "Class 13"
format: docx
execute:
  error: TRUE
---

# Plan
* Midterm postmortem
* Review of Merging / Model Diagnostics
* Group work
* Intro to Bivariate Regression (Lecture)
* Bivariate Regression in R (if time)

# Merging and Prediction Error Review

```{r}
library(tidyverse)

pres08 <- read.csv("./class_data/pres08.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/main/class_data/pres08.csv","pres08.csv")
#pres08 <- read.csv("pres08.csv")

polls08 <- read.csv("./class_data/polls08.csv") 

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/main/class_data/polls08.csv","polls08.csv")
#polls08 <- read.csv("polls08.csv")

pres08 <- pres08 |> mutate(act.margin= Obama - McCain) 
polls08 <- polls08 |> mutate(poll.margin = Obama - McCain)

```

Let's filter to the most recent poll in each state

```{r}
polls08 |> group_by(state) |> slice(1)
```

We can use |> arrange() to sort data. Let's use this to get the earliest poll in each state

```{r}

```

Now let's save just the columns we need:

```{r}
recent <- polls08 |> group_by(state) |> slice(1) |> select(state,poll.margin)
```

Now let's merge the dataset from above to the actual election results, and save it as a new dataset

```{r}
merged <- pres08 |> left_join(recent,by="state")
```

Create a new variable that measures the error

```{r}
merged <- merged %>% mutate(error=act.margin-poll.margin)

```

Evaluate the quality of our predictions, using RMSE.

```{r}

sqrt(mean(merged$error^2))

```

# Group Work

Modify the code above to average across all polls conducted _within the last two weeks_ before the election. Does the accuracy increase or decrease compared to looking at the most recent poll?
  
Coding Steps:
1. Use filter() to subset the initial polls dataset to include only polls with days.until <= 14
2. Use |> group_by()|> summarize(average14=xxxx) on the filtered dataset from #1, to obtain the _average_ poll margin in each state. Save the dataset with a new name.
3. Merge this new dataset to the actual election results (pres08) using left_join()
4. Create a new variable in the merged dataset measuring the difference between the averaged polls and the actual results
5. Compare the RMSE to the model we generated in class

```{r}





```


# Bivariate Regression (lecture)

## Basic Regression in R 

We are going to examine data on support for the Nazi party in Germany in 1933.

```{r}
library(tidyverse)
fascism <- read.csv("./class_data/fascism.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/class_data/fascism.csv","fascism.csv")
#fascism <- read_csv("fascism.csv")

fascism$nazi_voteshare <- fascism$nazivote/fascism$nvoter
fascism$unemployment_rate <- fascism$nunemployed/fascism$nvoter
```

Fitting regressions in R is straightforward: we can use the function lm()

```{r}
y <- fascism$nazi_voteshare
x <- 
  

```

To visualize, we can build a scatter plot, and superimpose the regression line on the data using abline():

```{r}
plot(x,y)

```

# Remember to complete the survey (link posted in Slack)

# Resources

Files > Book Chapters > Data Analysis, CH 4

## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand (leave blank if nothing)?

