---
title: "Class 1"
format: docx
---

Plan:
-   Class information
-   Project management with R / git
-   Learning Markdown

# Markdown

## A simple markup language

Markdown allows us to format text. Markdown can be compiled directly into a PDF, a Word file or an HTML file.

For this class we only care about some basic formatting:

**Bolding text**
*Italicizing text*

The **quick** brown fox jumps over the *startled* gopher

# Creating a larger header

## Creating a second level header

## Why Notebooks?

R Notebooks allow us to write text, write code and generate results within a single file. In this class all in-class assignments and homework assignments will be submitted using Markdown Notebooks.

## How to write code within markdown

When you see a white background you are writing text. You can use Markdown here. However we will also be interested in writing code to explore and summarize data. Code is placed in a *chunk.* Chunks start and end with three grave accents (the same key as the tilde): chunks are interactive and allow to you run code. When you execute code within the notebook, the results appear beneath the code. To run a chunk, press the *Run* button (play button) within the chunk.

Multiply 5 and 10:

```{r}
# NOTE: we cannot use markdown in code chunks.  Here # means that we are creating a comment (notes that are not executable code).  We are not creating a header.
```

Multiply 4 and 6, then square it

```{r}

```

R also contains *functions*. For instance, the function sqrt() calculates a square root:

```{r}

```


# Exercise 1

Expanding this notebook, complete the following tasks:

1)  Create two new code chunks (using the green 'C' button in the top right).

2)  In the first chunk, find the sum of 75, 83, 93, 34, 75 and 81 using math operators (+). Then try to find the sum using the function *sum()*. You can look up how to use the function *sum()* by typing ?sum in the console (bottom of your screen).

3)  In the second chunk divide 75 by 5, then add 43 and then multiply 6 (use parentheses)

# Exercise 2

Using markdown, answer the following questions:

1)  Where are you from?
2)  What is your class year?
3)  What is your major/planned major?
3)  Have you taken a statistics or a programming class before?
4)  What is your favorite book or movie?

# Prepare for submission

You will submit all assignments and problem sets as Markdown files. To create a markdown file, click the Render icon. 

# Check-in

1.  On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class:
2.  What was the easiest thing to understand?
3.  What was the most difficult thing to understand?

# Optional Resources

Canvas->Files->"Data Analysis Chapter 1"
  - Covers using R and basic operations
  