---
title: "Class 3"
format: docx
execute:
  error: TRUE
---

# Plan
* Review
* Learning R: More on manipulating data frames
* Learning R: Handling missing data
* Learning R: Practice with data frames (group work)

# Review: Notes on variable creation

```{r}
# Uncomment one of the options below to load your file.

# Git
#turnout <- read_csv("class_data/turnout.csv")

# Non-git
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/main/class_data/turnout.csv","turnout.csv")
#turnout <- read.csv("turnout.csv")

```

Using the *$* symbol to reference specific columns in the dataset, create a new variable in the dataframe called **VEP_turnout** that contains the turnout rate for each year. To do so, divide the total votes by the voting eligible population.

```{r}

# <- turnout$total_votes / turnout$VEP

```

Using your new variable, what is the average turnout rate in the data? What is the median turnout rate? You may have to look up the functions for this

```{r}

```

Create a new variable named *anes_vep_diff* that computes the difference between the actual turnout rate (VEP_turnout) and ANES estimates of the turnout rate for each year.

```{r}
#turnout$anes_vep_diff <- 
```


# Manipulating Data frames

# Creating variables with tidyverse

Another way to do create a variable without writing out turnout$ over again is to use the %>% mutate() function. mutate() is also a shortcut when creating many variables.

```{r}

turnout %>% mutate()

```

To save the changes made by %>% mutate(), we need to overwrite the original data frame

```{r}

```


## Removing specific columns

If we want to get rid of a column, we can put "-" in front of the column name when running select(). Let's return the data frame, removing the *VEP* column.

```{r}

turnout %>% select()

```


## More filtering

Let's filter the dataset to all years after 1990. To do this, we have to use the > operator within filter().

```{r}
turnout %>% filter()
```

We can also use >= or <=

```{r}
turnout %>% filter()
```

If we want to exclude something, we use '!=' Let's exclude the 1984 observation.

```{r}
turnout %>% filter()
```


## Handling missing data

R uses the value _NA_ to indicate missing data. Missing data will cause many functions to fail.

Let's place some missing data in our data frame.

```{r}
turnout$VEP_turnout[1] <- NA
```

Now, let's take the average of *VEP_turnout* using mean():

```{r}

```

For most functions, we can pass an additional argument to instruct R to ignore NAs, called na.rm=T

```{r}

```

## Summary statistics and chaining

Recall that the last group work asked you to separately find the average difference in turnout for midterm and presidential elections. The most efficient way to do this was to save a new dataset:

```{r}

midterms <- turnout |> filter(type=="Midterm")
mean(midterms$anes_vep_diff)

presi <- turnout |> filter(type=="Presidential")
mean(presi$anes_vep_diff)

```

We can also use |> summarize() to return summary statistics. To use summarize(), we place the the summary statistic function inside the parentheses: %>% summarize(function(column)) 

```{r}

midterms %>% summarize()

```

This is not super efficient. But with tidyverse, we can also do multiple pipe operations at once. They always execute from left to right:

dataset |> function1 |> function2

Let's do the filtering and the summarize in one step:

```{r}

```

How do we handle NAs using summarize()?

```{r}
turnout$anes_vep_diff[1] <- NA

```



# Group Work
Due on Sunday

Prior to the Syrian refugee crisis in 2015, many European political parties were already capitalizing on the perception that immigration to Europe had reached unsustainable levels. To investigate this trend, we will draw on demographic data provided by Eurostat. 

```{r}
# Git
foreignpop <-  read.csv("./class_data/foreignpopulation.csv")

# Non-git
download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/class_data/foreignpopulation.csv","foreignpopulation.csv")
foreignpop <- read.csv("foreignpopulation.csv")

```

The dataset contains 6 columns:

country                 Country name
foreign_2011            Number of foreign-born residents in 2011
foreign_2015            Number of foreign-born residents in 2015
population_2011         Total number of residents in 2011
population_2015         Total number of residents in 2015
western                 1 if the country is traditionally in Western Europe (non-communist), 0 otherwise

1. Use |> mutate() to create two new variables within the data frame: pctforeign_2011 and pctforeign_2015. These variables should contain the percentage of each country's population that is foreign-born in each year. *Remember to overwrite the original data frame to save the results.*

```{r}

```

2. Create a new variable in the data frame, delta_pctforeign, that measures the percentage point change in each country's foreign born population between 2011 and 2015. 

```{r}

```

3. What is the average change, across all countries, in the percentage of foreign population between 2011 and 2015? What is the median change? 

```{r}

```

4. Let's exclude countries with very small populations. Filter the data frame to only include countries with more than 5 million residents in 2015. Then recalculate the answer to #3. You can either save a new version of the dataframe, or use chaining.

```{r}

```

5. Let's look at only those countries in Western Europe. Using the original version of the dataframe, filter the data and recalculate your answer to #3. You can either save a new version of the dataframe, or use chaining.

```{r}

```

6. Lastly, let's look at only countries NOT in Western Europe. What is the average change in the percentage of foreign population between 2011 and 2015?

```{r}

```

## Optional Resources

Review:
https://datacarpentry.github.io/R-ecology-lesson/working-with-data.html 
  - Revoiew basic tidyverse commands we've covered so far
  
Prepare for Monday

https://jhudatascience.org/tidyversecourse/wrangle-data.html#filtering-data
  - Read about filtering for two conditions at once
  
https://jhudatascience.org/tidyversecourse/wrangle-data.html#grouping-data
  - Read about the group_by() function
 
Canvas --> Files/ Book Chapters/ Data Analysis, Chapter 2
  - Read about causality
  
## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
