---
title: "Class 5"
format: docx
execute:
  error: TRUE
---


# Plan
* Quiz
* Brief Lecture (Causality/Randomization)


# Brief Lecture: Causality / Randomization 


# Simulating the effects of randomization (demonstration only - do *not* need to know this code)

Initialize

```{r}
nrows <- 10000
age <- round(rnorm(10000,mean=40,sd=10))
college <- rbinom(10000,1,.38)
hair <- c(rep("brown",4000),rep("blonde",1000),rep("red",250),rep("black",4750))
handshake10k <- data.frame(age,college,hair)
```

Randomize

```{r}
handshake10k$treatment <- 0
randomsample <- sample(1:nrow(handshake10k),5000)
handshake10k$treatment[randomsample] <- 1
```

Let's check how accurate randomization was. Compare the mean value of age and college between the groups:

```{r}


```

Visualize Law of Large Numbers via flipping a fair coin

```{r}
n <- 50000
x <- sample(0:1,n, repl=T)
s <- cumsum(x)
r <- s/(1:n)
plot(r, ylim=c(0, 1), type="l")
lines(c(0,n), c(.50,.50),col="red")
round(cbind(x,s,r), 5)[1:10,]; r[n]
```


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?