---
title: "Class 21"
format: docx
execute:
  error: TRUE
---

# Plan
- qnorm() review
- Group work: Confidence Intervals
- t-statistics (lecture)
- t-statistics in R

# Review: qnorm()

```{r}
# Normal distribution (don't need to know this code)
x <- seq(-4, 4, length=100)
hx <- dnorm(x)
plot(x,hx,lty=1,type="l")

# 90% confidence interval

#abline(v=qnorm(),col="burlywood")
#abline(v=qnorm(),col="burlywood")

# 95% confidence interval

#abline(v=qnorm(),col="thistle")
#abline(v=qnorm(),col="thistle")

# 99% confidence interval

#abline(v=qnorm(),col="darkgoldenrod")
#abline(v=qnorm(),col="darkgoldenrod")

# 60% confidence interval

#abline(v=qnorm(),col="dodgerblue")
#abline(v=qnorm(),col="dodgerblue")


```

# Using regression for standard error / confidence interval

```{r}
library(tidyverse)
chechen <- read.csv("./class_data/chechen.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/class_data/chechen.csv","chechen.csv")
#chechen <- read.csv("chechen.csv")

se <- 0.4165658
diff_mean <- -0.553459

# 95% confidence interval
#qnorm((1-.95)/2)

diff_mean -1.959964*se
diff_mean +1.959964*se

```

For a _binary treatment indicator_, standard errors for treatment effects can be quickly calculated using linear regression. Let's regress fire on postattack 

```{r}

# lm(,data=chechen)

```

We can use _confint(,level=)_ to quickly calculate confidence intervals for regression coefficients.

```{r}
# confint(,level=)
```


# Group Work: Confidence Intervals

Let's re-evaluate the study on racial bias in hiring. 

```{r}
resume <- read.csv("./class_data/resume.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/main/class_data/resume.csv","resume.csv")
#resume <- read.csv("resume.csv")
```

In this experiment, identical resumes were sent out, except for the signaled race and gender, which was randomly assigned. We are interested in whether a resume received a call back from an employer or not.

Variables:
race    Signaled race on resume
sex     Signaled gender on resume
call    A binary variable indicating a call back (1) or not (0) 

1.1 First, calculate the average treatment effect:

Avg (Callback for White Resumes) - Avg (Callback for Black Resumes) 

```{r}

```

1.2 Now calculate the standard error for this difference in means. You can either use this formula:
sqrt( (sd_treatment^2 / n_treatment) + (sd_control^2 / n_control)), or a regression.

```{r}
n_treatment <- 
n_control <-
sd_treatment <- 
sd_control <-
  
# lm(data=resume,)  
```

1.3 Calculate a 90% confidence interval for the difference in means. 

```{r}

```

1.4 Can we conclude based on 1.3 that there is likely racial bias in hiring, with 90% confidence?



1.5: Would this change with a 99% confidence interval (critical value = 2.58)?

```{r}

```

1.6 Conceptual question:

The 90% confidence interval for the average of Group A is -0.04 to 0.07
The 90% confidence interval for the average of Group B is -0.03 to 0.17

Can we conclude that B > A?



# Lecture: t-distributions and t-statistics

## Demonstration in R

Compare t to normal (don't need to know this code)
```{r}
# Do not need to know this code
n <- samplesize <- 5
x <- seq(-4, 4, length=100)
hx <- dnorm(x)
plot(x,hx,lty=1,type="l",col="blue")
lines(x, dt(x,n-1), lwd=2, col="red")
```

For smaller sample sizes, we can use _qt()_ to obtain critical values and confidence intervals. This is the equivalent of _qnorm_ for the t-distribution. qt() requires us to state the degrees of freedom.

Let's obtain critical values for a 95% confidence interval, with a sample size of 10

```{r}
# qt( level, df)
```

We can use qt() all the time to be conservative, but if the sample size is large it will be similar to qnorm(). Let's use qt() for the resume example, constructing a 90% confidence interval

```{r}

```


# t-statistics

Manual calculation of a t-statistic:

Mean:          .19
Sample SD:   .3294
Sample Size:    20

```{r}

# First calculate the standard error

# Then calculate the t-statistic

```

Manual calculation of a t-statistic for a difference in means:

```{r}

result <- chechen %>% group_by(fire) %>%
summarize(out = mean(postattack), n = n(), variance = var(postattack))

diff <- result$out[2] - result$out[1]
se <- sqrt(result$variance[2]/result$n[2] + result$variance[1]/result$n[1])

# Calculate t-statistic (ate/se)

# Alternative: regression

```

What does a t-statistic of -1.33 mean?

```{r}
# Do not need to know this code
n  <- 318
x <- seq(-4, 4, length=100)
plot(x,dt(x,n-1),lty=1,type="l")
abline(v=-1.33)
```

The _t.test_ command can also rapidly generate t-statistics. It is more generalizable that regression, since we can do it for a single variable (rather than a relationship between two variables)

## One sample t-test (for a single distribution):

Let’s test whether the callback rate for black resumes is different from 0, at the 99% level.

```{r}
resume_b <- resume |> filter(race == "black")

#t.test(, conf.level)
```


We can also test whether it is different from another value (such as 2%), using the ,mu option:

```{r}
#t.test(, conf.level, mu=)
```


## Only If time: two sample t-test  (t-test for difference in means):

Let's test whether the difference in callback rates for black and white resumes is different from each other (statistically significant), at the 95% level

```{r}

# t.test(outcome1,outcome2,var.equal=F)

```

We can also use t-tests to check for accurate randomization. Let's check that female was independently randomized between the white and black resumes, at the 95% level:

```{r}


```

We can also use regression to obtain t-statistics. Note that regression always tests the null hypothesis that the true coefficient is 0.

```{r}

result <- lm(,data=resume)
summary(result)

```

## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?