---
title: "Class 20"
format: docx
execute:
  error: TRUE
---

# Plan

* Brief quiz discussion/review
* Discussion of group problem set
* Recap
* Group work: Error in ATE
* Confidence Intervals (lecture)
* Group work: Confidence intervals
* Standard error in linear regression

# Recap

# Group Work: Error in Average Treatment Effects

Let's work with the Chechen artillery fire natural experiment. First, obtain the difference in means between the treated and control group for the _postattack_ variable.

Treated group: (fire: 1)
Control group: (fire: 0)

```{r}
library(tidyverse)
chechen <- read.csv("./class_data/chechen.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/class_data/chechen.csv","chechen.csv")
#chechen <- read.csv("chechen.csv")

# Reminder: we can use group_by |> summarize() for this

```

Estimate the standard error for this difference in means, by obtaining the standard deviation and sample size of _each_ distribution. Remember that you can use table() to count; you can also use n() inside summarize(). To obtain the standard deviations for each group, use sd(outcome) on a filtered dataset, or use sd(outcome) within summarize()

```{r}

# Sample Size
n_treatment <- 
n_control <- 
  
# Standard deviation
sd_treatment <-
sd_control <-  	

# Combine using formula
#sqrt( sd_treatment^2 / n_treatment + sd_treatment^2 / n_control )
  
# Answer should match: 0.4165658
```

# Confidence Intervals (brief lecture)

# Confidence Intervals in R

The actual support for Trump in this particular state is 50%:

```{r}
state <- c(rep(1,50000),rep(0,50000))
```

How precise is our estimate of Trump's support, given the sample size and the dispersion in the outcome?

```{r}
# Take a single sample of size 1000
ss <- 1000
poll <- sample(state,ss,replace=F)

# Calculate the mean and standard error
p_mean <- 
p_se <- 
  
```

Since this is a random sample, let's take the information about the standard error to make an inference about how repeated measurements of the same population could vary (how far off could we be, still drawing from the 'correct population)?

Let's asssume a normal distribution, and use the 95% critical values

```{r}

```

Is the 'true value' in this confidence interval? (With our choice of critical values, roughly 5% of you will have a confidence interval that does not contain the true value)

Instead of memorizing 1.96, can use the _qnorm()_ function to get exact critical values from the normal distribution.


```{r}

```


# Group work: Confidence intervals  

Let's build on the previous group work, and construct confidence intervals for our Chechnya estimates.

Obtain 90%, 95%, and 99% confidence intervals for the Chechen difference in means you calculated above. You'll need the difference in means and the standard error. Use _qnorm()_ to obtain the exact critical values.

```{r}

```

Interpret the 90% confidence interval using appropriate language. 

[sentence here]



# With full class: (Only if time)

For a _binary treatment indicator_, standard errors can be quickly calculated using linear regression. Let's regress fire on postattack 

```{r}

```

We can use _confint()_ to quickly calculate confidence intervals for regression coefficients.

```{r}

```


## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?