---
title: "Class 17"
format: docx
execute:
  error: TRUE
---

# Plan
* Review on Creating Predictive Models
* Regression and ATE (lecture)
* Group work
* Limits of multivariate models 


# Review on Creating Predictive Models

Let's create a predictive model with the latest poll from 2008 and the percent of Black voters, and use it to predict outcomes.

To do this, we need the following:

*Training Dataset:*
Actual Margin in 2008 (y)
Poll Margin in 2008 (x)
% Black  in 2008 (x)

*Test Dataset:*
Poll Margin in 2020 (x)
% Black in 2020 (x)


```{r}
# All margins defined as Democrat - Republican

library(tidyverse)

polls08 <- read.csv("./problem_sets/data/polls08.csv")
polls12 <- read.csv("./problem_sets/data/polls12.csv")
polls16 <- read.csv("./problem_sets/data/polls16.csv")
polls20 <- read.csv("./problem_sets/data/polls20.csv")
pres08 <- read.csv("./problem_sets/data/pres08.csv")
pres12 <- read.csv("./problem_sets/data/pres12.csv")
pres16 <- read.csv("./problem_sets/data/pres16.csv")
demographics <- read.csv("./problem_sets/data/dem_race.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/polls08.csv","polls08.csv")
#polls08 <- read.csv("polls08.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/polls12.csv","polls12.csv")
#polls12 <- read.csv("polls12.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/polls16.csv","polls16.csv")
#polls16 <- read.csv("polls16.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/polls20.csv","polls20.csv")
#polls20 <- read.csv("polls20.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/pres08.csv","pres08.csv")
#pres08 <- read.csv("pres08.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/pres12.csv","pres12.csv")
#pres12 <- read.csv("pres12.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/pres16.csv","pres16.csv")
#pres16 <- read.csv("pres16.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/dem_race.csv","demographics_full.csv")
#demographics <- read.csv("demographics_full.csv")


pres08$actual.margin <- pres08$Obama - pres08$McCain
polls08$days.until <- as.Date("2008-11-04") - as.Date(polls08$middate, "%m/%d/%y")
polls08$margin <- polls08$Obama - polls08$McCain
polls16$margin <- polls16$clinton - polls16$trump
polls16$days.until <- as.Date("2016-11-08") - as.Date(polls16$date, "%m/%d/%y")
polls20$margin <- polls20$biden - polls20$trump
pres16$actual.margin <- pres16$clinton - pres16$trump
polls12$margin <- polls12$Obama - polls12$Romney
pres12$actual.margin <- pres12$Obama - pres12$Romney

```

```{r}

### CREATE THE TRAINING DATASET AND FIT THE MODEL

# Get most recent poll in 2008, by state
polls08_latest <- polls08 |> group_by(state)  |> 
  arrange(days.until)  |>  slice(1)  |> 
  select()

# Find the % black in 2008, by state

demographics08 <- demographics |> filter() |> mutate()

# Merge the two independent variables together

training <- polls08_latest  |> left_join()
  
# Merge in the outcomes for 2008

training <- training |>  left_join()
  
# Fit the model
result <- lm(data=training,)
summary(result)

```


```{r}

### CROSS-VALIDATE ON 2016 (predict)

# Get most recent poll in 2016, by state
polls16_latest <- polls16 |> group_by(state) |>  
  arrange(days.until) |>  slice(1) |> 
  select()

# Find the % black in 2016, by state
demographics_20 <- demographics |> filter() |> mutate()

# Merge the two together
xvalidate <- polls16_latest |> left_join()
  
# Create a test dataframe
predict_frame <- data.frame()

# Predict
predictions <- predict(result,newdata=predict_frame)
xv.out.df <- data.frame(state=xvalidate$state,predict=predictions)
xv.out.df


### Diagnose

# Merge in actual outcomes from 2016

xv.out.df <- xv.out.df 

# Calculate error

# RMSE
sqrt(mean(ERRORˆ2))

# Plot

x<- 
y<- 
  
plot(type="n",x =  x , y = y, xlab = "Predicted Results", ylab = "Actual Election Results")
text(x = x  , y = y, 
     labels = , 
     col = "blue",
     cex=.5)

# Diagnosis Lines
abline(0, 1, lty = "dashed")
abline(v=0)
abline(h=0)

```


```{r}

### CREATE THE TEST DATASET AND PREDICT ON 2020

# Get most recent poll in 2020, by state
polls20_latest <- polls20  |>  group_by(state)  |>  
  arrange(days.until)  |> slice(1)  |> 
  select()

# Find the % black , by state

demographics_20 <- demographics |> filter() |> mutate()

# Merge the two together
test <- polls20_latest |> left_join()
  
# Create a test dataframe
predict_frame <- data.frame()

# Predict
out <- predict(result,newdata=predict_frame)

out.df <- data.frame(state=test$state,predict=out)
out.df


```


# Brief Lecture: Regression and Causation


# Group work

Let's examine a simple case first, with 1 treatment. 

```{r}
resume <- read.csv("./class_data/resume.csv")
#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/main/class_data/resume.csv","resume.csv")
#resume<- read.csv("resume.csv")
```

Calculate the ATE for the racial discrimination resume experiment using group_by |> summarize().

The outcome variable is *call* (1 or 0), and the treatment variable is *race* (black or white)

```{r}

```

Now do it with a regression

```{r}

```


Let's examine the Kenyan savings experiment again:

```{r}
kenya <- read.csv("./class_data/rosca2.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/main/class_data/rosca2.csv","kenya.csv")
#kenya<- read.csv("kenya.csv")

kenya <- subset(kenya,has_followup2 == 1)
```

fol2_amtinvest            The outcome (investment in Kenyan shillings)
encouragement             1 if the person is in the control group, 0 otherwise
locked_box                1 if the person is in the locked box group, 0 otherwise
safe_box                  1 if the person is in the safe_box group, 0 otherwise
bg_b1_age                 Age of participant
bg_married                Whether participant is married (1) or not (0)
bg_female                 Whether participant is female (1) or not (0)

Calculate the ATE of the locked_box and safe_box groups, using a single multivariate regression model.

```{r}


```

Now add the background characteristics of participants to the model. Do the ATEs change? What might this tell you about the quality of the randomization? Interpret the coefficient on Locked Box.

```{r}


```


# Demonstration: limits on adding new predictors to multivariate models.

*Degrees of freedom:*
This corresponds to our number of observations (n), minus 1, minus the number of independent variables in our model. Degrees of freedom must always be positive to return cofficients. If degrees of freedom is too small, we also become more uncertain about whether our coefficients are accurate.

*Collinearity:*

If independent variables are perfectly correlated, we cannot estimate both terms. Very high correlation is also problematic for accurate coefficients (difficult to disentangle direct and indirect effects)

```{r}

# Number of observations

kenya_subset <- kenya %>% slice_sample(5)
result <- lm(fol2_amtinvest ~ safe_box + locked_box + bg_b1_age + bg_married + bg_female, data = kenya_subset)
summary(result)

# Collinearity

kenya <- kenya %>% mutate(bg_unmarried = ifelse(bg_married==0,1,0))
kenya$bg_b1_age_18 = kenya$bg_b1_age - 18

result <- lm(fol2_amtinvest ~ safe_box + locked_box + bg_b1_age + bg_married  + bg_female, data = kenya)
summary(result)
```

## Check-in

1. On a scale ranging between 1 (Too Hard) and 10 (Too Easy), how was today's class: 
2. What was the easiest thing to understand?
3. What was the most difficult thing to understand?
