---
title: "Problem Set 3"
format: docx
execute:
  error: TRUE
---


## Municipal Mergers  (Difference-in-Differences)

For many years, researchers have argued that small-sized municipalities create the potential for welfare gains because public sectors can be better tailored to local preferences. However, a counterargument holds that the large-sized municipalities are more cost-effective due to economies of scale.

If scale effects exist in municipalities, we should, all things equal, observe a negative correlation between municipality size and per capita administrative costs.

The Danish structural reform in late 2007 provides an unique opportunity. The reform was quick, based on objective factors, and mergedmany municipalities within six months. This created a variation between merged (treated) and not merged (control).

Note that amalgamation results in a change in the unit of analysis. The control group (those that did not experience amalgamation) has the same municipalities before and after the structural reform.  However, the units in the treatment group change due to merging. In this exercise, we designate merged municipalities as treated units and calculate their outcomes before amalgamation as an average. For example, suppose a municipality A in this dataset is a result of merging the municipalities B, C, and D. During the pre-treatment period (2005-2007), the outcome `costs` is the average of the outcomes from B, C, and D.

------------------------------------------------------------------------------
 Name                Description
 ------------------- ---------------------------------------------------------
 `year`               Year. (pre-treatment period is 2005-2007, post-treatment
                        period is 2008-2011)

 `costs`              The outcome variable. Administrative costs per capita (1000s of Danish Kroner)
                     
 `treatment`          The municipality is a result of amalgamation (1, treated group)
                        or not (0, control group).
 
 `region`             Region ID (1-5)  
 
 `municipality_id`    Municipality ID (there are 98 unique municipalities)
 
------------------------------------------------------------------------------

```{r}
library(tidyverse)
admin_costs <- read.csv("./data/admin_costs.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/admin_costs.csv","admin_costs.csv")
#admin_costs <- read.csv("admin_costs.csv")

```

### Question 1

2.1 Calculate the difference-in-differences (DiD) estimate, using the years 2007 (pre) and 2008 (post). Contextualize the effect size.

```{r}
                      
```

*short answer here*

2.2 Calculate the average administrative costs per capita in each year, separately for the treated and control groups. Then plot the data with a line plot. Adjust the y limits on the y-axis using ylim() so that it displays values between 4500 and 6500 kroner.

Hint: Use plot(x,type=“l”,ylim=c()), where x is the vector. You can add the second vector using the lines()
function. Remember to color each line differently


```{r}


```


2.3 Calculate the difference-in-differences (DiD) estimate for the years 2006 (pre) and 2007 (post). 

```{r}


```


2.4 Do the results from 2.3 increase or decrease your confidence that the relationship is causal? Why?

*short answer here*

# Voting in the United Nations General Assembly (Correlation)

Like legislators in the US Congress, the member states of the United Nations (UN) are politically divided on many issues such as trade, nuclear disarmament, and human rights. During the Cold War, countries in the UN General Assembly tended to split into two factions: one led by the capitalist United States and the other by the communist Soviet Union. In this exercise, we will analyze how states' ideological positions, as captured by their votes on UN resolutions, have changed since the fall of communism.

In the analysis that follows, we measure state preferences in two ways. First, we can use the proportion of votes by each country that coincide with votes on the same issue cast by the two major Cold War powers: the United States and the Soviet Union. For example, if a country voted for ten resolutions in 1992, and if its vote matched the United States’s vote on exactly six of these resolutions, the variable PctAgreeUS in 1992 would equal 60 for this country. 

Second, we can also measure state preferences in terms of numerical ideal points. These _ideal points_ capture what international relations scholars have called countries’ liberalism on issues such as political freedom, democratization, and financial liberalization. 

## Load the data

```{r}
unvotes <- read.csv("./data/unvoting.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/unvoting.csv","unvoting.csv")
#unvotes <- read.csv("unvoting.csv")

```

| Name           | Description                                                  |
|----------------|--------------------------------------------------------------|
| CountryName    | The name of the country                                      |
| CountryAbb     | The abbreviated name of the country                          |
| idealpoint     | Estimated ideal point                                        |
| Year           | The year for which the ideal point is estimated              |
| PctAgreeUS     | The proportion of votes that match votes cast by the US      |
| PctAgreeRUSSIA | The proportion of votes that match votes cast by Russia/USSR |


2.1 What is the correlation between _idealpoint_ and _PctAgreeUS_? Between _idealpoint_ and _PctAgreeRUSSIA_ ? 

```{r}

```

2.2. Let's examine the distribution of ideal points in 1950, at the start of the cold war, versus in 1985, during the height of the cold war. To do so, create a boxplot that puts each distribution side by side. Use the following syntax: boxplot(distribution1,distribution2)

```{r}

```

2.3 How do the distributions of ideal points differ between 1950 and 1984? 

*Short sentences here*


2.4. Now let's turn to changes over time. Create a line plot that measures the average percent agreement with the US in each year. Set the limits of the y axis to range between 0 and 1. The code for this is in class10.qmd; you'll have to adapt it slightly.

```{r}

```

2.5. Add the average percent agreement with Russia as a red line for comparison  

Note: You will need to copy your code from 2.4 into this chunk, generate a vector for Russia, and then use lines(x, col="red")

```{r}

```

2.6. Were countries moving closer to the US or to Russia during the cold war (before 1991)? What about after Putin came to power (after 1999)? 

*short answer here*


2.7 One problem with using the proportion of votes that agree with the US or Russia as a measure of state preferences is that the ideological positions of the US and Russia might also change over time. This makes it difficult to know if the US, Russia or both changed ideological positions. Investigate this issue by plotting ideal points for the US and Russia over time on the same graph. 

You will need to use the *CountryAbb* variable to select ideal points for only the US or Russia, respectively. Then group by year and use the summarize() function to generate the vectors for the line plots. 

```{r}

```

2.8 To truly understand the pattern, we will need to plot movement among all countries excluding Russia and the US.  Compute the mean ideal point for all countries (excluding the US and Russia) each year and add this to the plot (you will need to copy your code from 2.7 to the chunk below).

```{r}

```

2.9 How might the results of this analysis modify (or not) your interpretation of the previous analysis? 

*short answer here*
