---
title: "Problem Set 4"
format: docx
execute:
  error: TRUE
---

# Poll Averages and Election Predictions 

We will evaluate data from the 2016 Presidential Election, which featured Trump vs. Clinton. The election date was the 8th of November, 2016.

```{r}
library(tidyverse)

polls16 <- read.csv("./data/polls16.csv") 
pres_16 <- read.csv("./data/pres_16.csv")


#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/polls16.csv","polls16.csv")
#polls16 <- read.csv("polls16.csv")


#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/pres_16.csv","pres_16.csv")
#pres_16 <- read.csv("pres_16.csv")

```


1.1 In the polls16 dataset, create a variable which measures the polling margin, which we will define as (Trump-Clinton). In the pres16 dataset, create a variable that measures the actual voting margin.

```{r}

```

1.2 Filter the polls16 dataset to only include those that sample the population of likelyvoters (lv in the population column), with a sample size greater than 300. 

```{r}

```

1.3 Now, create a new version of the dataset that contains only polls conducted within 15 days of the last poll within each state. For instance, if the last poll in State A was conducted 3 days prior to the election, your dataset should include all polls conducted on days 18 through 3 for State A. If the last poll in State B was conducted 5 days prior to the election, your polls from this state should be drawn from days 5 through 20, etc.

_Note: There are 2 ways of doing this. Here are some pointers to get you started down the right path_

Option 1:
a. Create a new dataset called _latestpoll_ that contains the most recent poll in each state, and use select() to extract just the state abbreviation and the _days_until_ value for the most recent poll.
b. Use |> rename(lastpoll=days_until) on the results from #1 to rename _days_until_ into a new column called _lastpoll_
c. Merge the results of #2 back into the polls16 dataset, and filter to contain only polls within x days of _lastpoll_

Option 2:
Use  |> group_by |>  filter() , along with min() inside filter


```{r}

```

1.4 Start with your dataset containing all polls conducted within 15 days of the last poll, which you created in the previous question. Now, extract the average of the polls within each state, and save it as a new dataset called _avg_polls_ . Merge the _avg_polls_ and the _pres16_ dataset together into a new dataset called _merged.16_. 

```{r}

```

1.5 Using the _merged.16_ dataset, assess the relationship between the actual voting margin (y) and the polling margin (x) using a bivariate regression model. Obtain and interpret the regression coefficients. What do they tell you about the relationship of the polls to the actual election results in 2016? 

```{r}

```

*A few sentences here*

1.6 Let's examine whether other variables can potentially improve our prediction and reduce polling bias. The following dataset contains state-level demographic data:

```{r}
demographics <- read.csv("./data/demographics.csv")

#download.file("https://raw.githubusercontent.com/resonance1/govt10-s25/refs/heads/main/problem_sets/data/demographics.csv","demographics.csv")
#demographics <- read.csv("demographics.csv")
```

Filter the demographics dataset to 2016 only. Then create a new variable in the dataset called _under45_ that contains the proportion of residents who are 45 years or younger. Next, merge this dataset with your existing _merged.16_ dataset using left_join(): 

```{r}


```

1.7 Using the master dataset, fit two regressions: first, regress the actual voting margin in each state (y) on the `Male' variable (x), which measures the share of residents who are White (0 to 1 scale). Second, regress the actual voting margin in each state (y) on the _under45_ variable (x). Interpret the coefficients from each regression as well as the R^2.

```{r}

```

*A few sentences here*

1.8 Now fit a multivariate regression model using three independent variables: White, under45, and the 15-day poll margin according to the most recent state poll(s). The syntax is  _lm(y~ x1 + x2 + x3)_ , where y is the outcome, and x1, x2, and x3 are the independent variables. Interpret the R^2 in words. Then compare the R^2 between the multivariate model and the bivariate model calculated in 1.7. Does adding this additional information improve the predictive power of the model? 

```{r}


```

*2 sentences here*



Please submit using Canvas
